Miniproject - Algebraic Geometry

### 2021-12-12 Algebraic geometry/submission.tex
Fixed first part of proof of isomorphism in alg geo

    So in particular find for all $i_2, ..., i_d \leq m < i_1 \leq n$ that

@@

    For the other direction, let $v_1, ..., v_d$ be linearly independent.
    TODO
\end{proof}

@@

    Let $\omega = v_1 + ... + v_t$ for linearly independent decomposable vectors $v_i \in \extpow^2 V$.
    Then
    \begin{align*}
        0 =& \omega \wedge \omega = \sum_{i, j} v_i \wedge v_j = \sum_{i < j} (v_i \wedge v_j) + (v_j \wedge v_i) \\
        =& \sum_{i < j} 2(v_i \wedge v_j) = 2\sum_i v_i \wedge \Bigl( \sum_{j > i} v_j \Bigr)
    \end{align*}
    Here we used that the permutation $\bigl(1 \ 2d\bigr)\bigl(2 \ (2d - 1)\bigr)...\bigl(d \ (d + 1)\bigr) \in S_{2d}$ has always sign $1$ (since $d$ is even).

    Note that for any nonzero decomposable vector
    \begin{equation*}
        u_1 \wedge u_2 \in \Bigl( \extpow^2 \vspan\{v_2, ..., v_t\} \Bigr) \setminus \{0\}
    \end{equation*}
    find
    \begin{equation*}
        u_1, u_2 \in \vspan\{v_2, ..., v_t\}
    \end{equation*}
    In particular, we know that
    \begin{equation*}
        v_1 \wedge \Bigl( \sum_{j > i} v_j \Bigr) \in \extpow^2 \vspan\{v_2, ..., v_t\}
    \end{equation*}
    and so $v_1 \in \vspan\{v_2, ..., v_t\}$ unless $\sum_{j > i} v_j = 0$ by Lemma~\ref{prop:subspace_closed}.
    We assumed that the $v_i$ are linearly independent, so the former would give a contradiction.
    Hence $\sum_{j > i} v_j = 0$ and thus $t = 1$, i.e. $\omega = v_1$ is decomposable.
\end{proof}

@@

In this part, we want to consider the connection of external powers to the Grassmanian.
\begin{remark}[Question (a)]
    First of all, assume there are two bases $v_1, ..., v_d$ and $u_1, ..., u_d$ of a $d$-dimensional vector space $U$. 
    Then there exists a basis change matrix $A = (a_{ij}) \in \GL_d(k)$ with
    \begin{equation*}
        u_i = \sum_j a_{ij} v_j
    \end{equation*}
    So by Lemma~\ref{prop:linear_transform_extpow}, it follows that
    \begin{equation*}
        u_1 \wedge ... \wedge u_d = \det(A) (v_1 \wedge ... \wedge v_d)
    \end{equation*}
    As $v_1, ..., v_d$ resp. $u_1, ..., u_d$ are bases, they are linearly independent and in particular, we see that
    \begin{equation*}
        v_1 \wedge ... \wedge v_d \neq 0 \quad \text{and} \quad u_1 \wedge ... \wedge u_d \neq 0
    \end{equation*}
    by Lemma~\ref{prop:extpow_zero_iff_independent}.
    Hence they have well-defined images $[v_1 \wedge ... \wedge v_d]$ resp. $[u_1 \wedge ... \wedge u_d]$ in the projective space $\Proj(\extpow^d V)$.
    By the above, find
    \begin{equation*}
        [v_1 \wedge ... \wedge v_d] = [u_1 \wedge ... \wedge u_d]
    \end{equation*}
\end{remark}
This allows us to study the Grassmanian $\Gr(d, V)$ of a fixed vector space $V$.
\begin{definition}
    Define the map
    \begin{equation*}
        \phi: \Gr(d, V) \to \Proj(\extpow^d V), \quad \vspan\{v_1, ..., v_d\} \mapsto [v_1 \wedge ... \wedge v_d]
    \end{equation*}
    which is well-defined by Lemma~\ref{prop:linear_transform_extpow} as described above.
\end{definition}
\begin{lemma}[1a]
    We have
    \begin{equation*}
        \mathrm{im}\phi = D := \{ [v] \in \Proj(\extpow^d V) \ | \ \text{$v$ decomposable}\}
    \end{equation*}
\end{lemma}
\begin{proof}
    First of all, note that the set $D$ is well-defined, as $v$ is decomposable if and only if $\lambda v$ is decomposable, for all $\lambda \in k^*$.

@@

    This follows directly from Lemma~\ref{prop:extpow_zero_iff_independent}.
    Let $\Gr(d, n) := \Gr(d, k^n)$.
In the lecture, we considered an embedding of $\Gr(d, n)$ into projective space given by minors of the basis matrix.
This corresponds to the following definition.
    Define the map
        \rho: \Gr(d, n) &\to \Proj\Bigl(k^{\{1, ..., n\}^{(d)}}\Bigr) \cong \Proj^{{n \choose d} - 1}, \\
        \vspan\{v_1, ..., v_d\} &\mapsto \Bigl[ \det\left(\begin{matrix}
            v_{1i_1} & ... & v_{di_1} \\
            \vdots & \ddots & \vdots \\
            v_{1i_d} & ... & v_{di_d}
        \end{matrix}\right) \Bigr]_{\{i_1, ..., i_d\} \in \{1, ..., n\}^{(d)}}
    where $\{1, ..., n\}^{(d)} := \{ I \subset \{1, ..., n\} \ | \ \#I = d \}$ is the set of all $d$-element subsets of $\{1, ..., n\}$.
    \label{prop:isomorphism_extpow_det}
    There is a linear isomorphism
        f: \extpow^d k^n &\to k^{\{1, ..., n\}^{(d)}}, \\
        \sum_j v_1^{(j)} \wedge ... \wedge v_d^{(j)} &\mapsto \Bigl( \sum_j \det\left(\begin{matrix}
            v_{1i_1}^{(j)} & ... & v_{di_1}^{(j)} \\
            \vdots & \ddots & \vdots \\
            v_{1i_d}^{(j)} & ... & v_{di_d}^{(j)}
        \end{matrix}\right) \Bigr)_{\{i_1, ..., i_d\} \in \{1, ..., n\}^{(d)}}
    For vectors $v_1, ..., v_d$ and $I = \{i_1, ..., i_d\} \in \{1, ..., n\}^{(d)}$ write
        A_I(v_1, ..., v_d) := \left(\begin{matrix*}
            v_{1i_1} & ... & v_{di_1} \\
            \vdots & \ddots & \vdots \\
            v_{1i_d} & ... & v_{di_d}
        \end{matrix*}\right)
    First of all, we show that $f$ is well-defined.
    Note that the tensor product can be described as
        V^{\otimes d} := k^{V \times ...\times V} / \vspan\{ &\bigl(v_1 \otimes ... \otimes v_{i - 1} \otimes \bigl(v_i + v'_i\bigl) \otimes v_{i + 1} \otimes ... \otimes v_d\bigl) \\
        &- \bigl(v_1 \otimes ... \otimes v_d\bigl) - \bigl(v_1 \otimes ... \otimes v_{i - 1} \otimes v'_i \otimes v_{i + 1} \otimes ... \otimes v_d\bigl), \\
        & \bigl(v_1 \otimes ... \otimes v_{i - 1} \otimes \lambda v_i \otimes v_{i + 1} \otimes ... \otimes v_d\bigl) \\
        &- \lambda\bigl(v_1 \otimes ... \otimes v_{i - 1} \otimes v_i \otimes v_{i + 1} \otimes ... \otimes v_d\bigl) \ | \ i \leq d, v_j, v'_i \in V\}
    where $v_1 \otimes ... \otimes v_d := \chi_{(v_1, ..., v_d)}$.
    Hence the external power can be described as
        \extpow^d V := k^{V \times ... \times V} / \vspan\Bigl\{ &\bigl(v_1 \otimes ... \otimes v_{i - 1} \otimes \bigl(v_i + v'_i\bigl) \otimes v_{i + 1} \otimes ... \otimes v_d\bigl) \\
        &- \bigl(v_1 \otimes ... \otimes v_d\bigl) - \bigl(v_1 \otimes ... \otimes v_{i - 1} \otimes v'_i \otimes v_{i + 1} \otimes ... \otimes v_d\bigl), \\
        & \bigl(v_1 \otimes ... \otimes v_{i - 1} \otimes \lambda v_i \otimes v_{i + 1} \otimes ... \otimes v_d\bigl) \\
        &- \lambda\bigl(v_1 \otimes ... \otimes v_{i - 1} \otimes v_i \otimes v_{i + 1} \otimes ... \otimes v_d\bigl), \\
        &v_1 \otimes ... v_{j - 1} \otimes (v_j \otimes v_{j + 1} + v_{j + 1} \otimes v_j) \otimes v_{j + 1} \otimes ... \otimes v_d \\
        &\quad | \ i \leq d, \ j \leq d - 1, \ v_1, ..., v_d, v'_i \in V\Bigr\}
    So it suffices to show that for all $I \in \{1, ..., n\}^{(d)}$ and vectors $v_1, ..., v_d, v'_i \in V$
        \det(A_I(v_1, ..., v_i + v'_i, ..., v_d)) = \det(A_I(v_1, ..., v_d)) + \det(A_I(v_1, ..., v'_i, ..., v_d))
    and
        \det(A_I(v_1, ..., \lambda v_i, ..., v_d)) = \lambda\det(A_I(v_1, ..., v_d))
    and
        \det(A_I(v_1, ..., v_{j + 1}, v_j, ..., v_d)) = -\det(A_I(v_1, ..., v_d))
    However, these properties follow from the well-known properties of the determinant.
    In particular, $\det$ is linear in each column and swapping columns negates the determinant.
    It follows that $f$ is indeed well-defined.

>>>>

    So in particular find for all $i_2 < ... < i_d \leq m < i_1 \leq n$ that

@@

    For the other direction, let $v_1, ..., v_d$ be linearly independent.
    Clearly, we can extend them to a basis $v_1, ..., v_n$ of $V$.
    We then see that there is a matrix $A = (a_{ij}) \in \GL_n(k)$ with $v_i = \sum_j a_{ij} b_j$.
    So by Lemma~\ref{prop:linear_transform_extpow} have
    \begin{equation*}
        v_1 \wedge ... \wedge v_n = \underbrace{\det(A)}_{\neq 0} (b_1 \wedge ... \wedge b_n) \neq 0
    \end{equation*}
    So clearly $v_1 \wedge ... \wedge v_d \neq 0$ as $0 \wedge (v_{d + 1} \wedge ... \wedge v_n) = 0$ for all $v_{d + 1}, ..., v_n \in V$.
\end{proof}

@@

    Let $\omega = v_1 + ... + v_t$ for linearly independent decomposable vectors $v_i \in \extpow^2 V$.
    Then
    \begin{align*}
        0 =& \omega \wedge \omega = \sum_{i, j} v_i \wedge v_j = \sum_{i < j} (v_i \wedge v_j) + (v_j \wedge v_i) \\
        =& \sum_{i < j} 2(v_i \wedge v_j) = 2\sum_i v_i \wedge \Bigl( \sum_{j > i} v_j \Bigr)
    \end{align*}
    Here we used that the reversing permutation $\bigl(1 \ 2d\bigr)\bigl(2 \ (2d - 1)\bigr)...\bigl(d \ (d + 1)\bigr) \in S_{2d}$ has even parity, i.e. $\sgn(\cdot) = 1$ (since $d$ is even).

    Now note that
    \begin{equation*}
        v_1 \wedge \Bigl( \sum_{j > 1} v_j \Bigr) = 0 \in \extpow^2 \vspan\{v_2, ..., v_t\}
    \end{equation*}
    and so $v_1 \in \vspan\{v_2, ..., v_t\}$ unless $\sum_{j > 1} v_j = 0$ by Lemma~\ref{prop:subspace_closed}.
    We assumed that the $v_i$ are linearly independent, so the former would give a contradiction.
    Hence $\sum_{j > 1} v_j = 0$ and thus $t = 1$, i.e. $\omega = v_1$ is decomposable.
\end{proof}

@@

In this part, we want to consider the connection of exterior powers to the Grassmanian.
\begin{remark}[Question (a)]
    First of all, assume there are two bases $v_1, ..., v_d$ and $u_1, ..., u_d$ of a $d$-dimensional vector space $U$. 
    Then there exists a basis change matrix $A = (a_{ij}) \in \GL_d(k)$ with
    \begin{equation*}
        u_i = \sum_j a_{ij} v_j
    \end{equation*}
    So by Lemma~\ref{prop:linear_transform_extpow}, it follows that
    \begin{equation*}
        u_1 \wedge ... \wedge u_d = \det(A) (v_1 \wedge ... \wedge v_d)
    \end{equation*}
    As $v_1, ..., v_d$ resp. $u_1, ..., u_d$ are bases, they are linearly independent and in particular, we see that
    \begin{equation*}
        v_1 \wedge ... \wedge v_d \neq 0 \quad \text{and} \quad u_1 \wedge ... \wedge u_d \neq 0
    \end{equation*}
    by Lemma~\ref{prop:extpow_zero_iff_independent}.
    Hence they have well-defined images $[v_1 \wedge ... \wedge v_d]$ resp. $[u_1 \wedge ... \wedge u_d]$ in the projective space $\Proj(\extpow^d V)$.
    By the above, find
    \begin{equation*}
        [v_1 \wedge ... \wedge v_d] = [u_1 \wedge ... \wedge u_d]
    \end{equation*}
\end{remark}
This allows us to study the Grassmanian $\Gr(d, V)$ of a fixed vector space $V$.
\begin{definition}
    Define the map
    \begin{equation*}
        \phi: \Gr(d, V) \to \Proj(\extpow^d V), \quad \vspan\{v_1, ..., v_d\} \mapsto [v_1 \wedge ... \wedge v_d]
    \end{equation*}
    which is well-defined by Lemma~\ref{prop:linear_transform_extpow} as described above.
\end{definition}
\begin{lemma}[1a]
    We have
    \begin{equation*}
        \mathrm{im}\phi = D := \{ [v] \in \Proj(\extpow^d V) \ | \ \text{$v$ decomposable}\}
    \end{equation*}
\end{lemma}
\begin{proof}
    First of all, note that the set $D$ is well-defined, as $v$ is decomposable if and only if $\lambda v$ is decomposable, for all $\lambda \in k^*$.

@@

    This follows directly from Lemma~\ref{prop:extpow_zero_iff_independent}.
\end{proof}
\begin{definition}
    Let $\Gr(d, n) := \Gr(d, k^n)$.
\end{definition}
In the lecture, we considered an embedding of $\Gr(d, n)$ into projective space given by minors of the basis matrix.
This corresponds to the following definition.
\begin{definition}
    Define the map
    \begin{align*}
        \rho: \Gr(d, n) &\to \Proj\Bigl(k^{\{1, ..., n\}^{(d)}}\Bigr) \cong \Proj^{{n \choose d} - 1}, \\
        \vspan\{v_1, ..., v_d\} &\mapsto \Bigl[ \det\left(\begin{matrix}
            v_{1i_1} & ... & v_{di_1} \\
            \vdots & \ddots & \vdots \\
            v_{1i_d} & ... & v_{di_d}
        \end{matrix}\right) \Bigr]_{\{i_1, ..., i_d\} \in \{1, ..., n\}^{(d)}}
    \end{align*}
    where $\{1, ..., n\}^{(d)} := \{ I \subset \{1, ..., n\} \ | \ \#I = d \}$ is the set of all $d$-element subsets of $\{1, ..., n\}$.
\end{definition}
\begin{lemma}
    \label{prop:isomorphism_extpow_det}
    There is a linear isomorphism
    \begin{align*}
        f: \extpow^d k^n &\to k^{\{1, ..., n\}^{(d)}}, \\
        \sum_j v_1^{(j)} \wedge ... \wedge v_d^{(j)} &\mapsto \Bigl( \sum_j \det\left(\begin{matrix}
            v_{1i_1}^{(j)} & ... & v_{di_1}^{(j)} \\
            \vdots & \ddots & \vdots \\
            v_{1i_d}^{(j)} & ... & v_{di_d}^{(j)}
        \end{matrix}\right) \Bigr)_{\{i_1, ..., i_d\} \in \{1, ..., n\}^{(d)}}
    \end{align*}
\end{lemma}
\begin{proof}
    For vectors $v_1, ..., v_d$ and $I = \{i_1, ..., i_d\} \in \{1, ..., n\}^{(d)}$ write
    \begin{equation*}
        A_I(v_1, ..., v_d) := \left(\begin{matrix*}
            v_{1i_1} & ... & v_{di_1} \\
            \vdots & \ddots & \vdots \\
            v_{1i_d} & ... & v_{di_d}
        \end{matrix*}\right)
    \end{equation*}
    First of all, we show that $f$ is well-defined.
    Clearly $f$ is well-defined on the decomposable vectors, as the determinant is negated by swapping columns.

    Now we have to show that $f$ yields the same value for different sum representations of an element of $\extpow^d V$.
    The idea is just that the determinant is linear in each column, but the details will be slightly technical.

    Assume
    \begin{equation*}
        \sum_l v_1^{(l)} \wedge ... \wedge v_d^{(l)} = \sum_{i_1 < ... < i_d} \lambda_{i_1, ..., i_d} (e_{i_1} \wedge ... \wedge e_{i_d})
    \end{equation*}
    Then clearly
    \begin{align*}
        \sum_{i_1 < ... < i_d} \lambda_{i_1, ..., i_d} \det(A_I(e_{i_1}, ..., e_{i_d})) = \sum_{i_1 < ... < i_d} \lambda_{i_1, ..., i_d} \begin{cases}
            1 & \text{if $I = \{i_1, ..., i_d\}$} \\
            0 & \text{otherwise}
        \end{cases} = \lambda_I
    \end{align*}
    So it suffices to show that for all $I = \{i_1, ..., i_d\} \in \{1, ..., n\}^{(d)}$ have
    \begin{equation*}
        \sum_l \det(A_I(v_1^{(l)}, ..., v_d^{(l)})) = \lambda_I
    \end{equation*}
    Assume that $v_j^{(l)} = \sum_i \mu_{l, j, i} e_i$. Since the determinant is linear in each column, find
    \begin{align*}
        &\sum_l \det(A_I(v_1^{(l)}, ..., v_d^{(l)})) = \sum_l \sum_{1 \leq j_1, ..., j_d \leq n} \mu_{l, 1, j_1} \cdot ... \cdot \mu_{l, d, j_d} \det(A_I(e_{j_1}, ..., e_{j_d})) \\
        =& \sum_l \sum_{j_1, ..., j_d \ \text{distinct}} \mu_{l, 1, j_1} \cdot ... \cdot \mu_{l, d, j_d} \cdot \begin{cases}
            \pm 1 & \text{if $\{j_1, ..., j_d\} = I$} \\
            0 & \text{otherwise}
        \end{cases} \\
        =& \sum_l \sum_{\pi \in S_d} \sgn(\pi) \mu_{l, 1, i_{\pi(1)}} \cdot ... \cdot \mu_{l, d, i_{\pi(d)}}
    \end{align*}
    On the other hand, observe that also $\wedge$ is multilinear (i.e. linear in each component), so
    \begin{align*}
        &\sum_l v_1^{(l)} \wedge ... \wedge v_d^{(l)} = \sum_l \sum_{1 \leq j_1, ..., j_d \leq n} \mu_{l, 1, j_1} \cdot ... \cdot \mu_{l, d, j_d} \bigl( e_{i_1} \wedge ... \wedge e_{i_d} \bigr) \\
        =& \sum_l \sum_{j_1, ..., j_d \ \text{distinct}} \mu_{l, 1, j_1} \cdot ... \cdot \mu_{l, d, j_d} \bigl( e_{i_1} \wedge ... \wedge e_{i_d} \bigr) \\
        =& \sum_l \sum_{i_1 < ... < i_d} (e_{i_1} \wedge ... \wedge e_{i_d}) \sum_{\pi \in S_d} \sgn(\pi) \mu_{l, 1, i_{\pi(1)}} \cdot ... \cdot \mu_{l, d, i_{\pi(d)}}
    \end{align*}
    By Lemma~\ref{prop:basis_product_independence} the $e_{i_1} \wedge ... \wedge e_{i_d}$ are a basis, and so we must already have that
    \begin{equation*}
        \lambda_{i_1, ..., i_d} = \sum_l \sum_{\pi \in S_d} \sgn(\pi) \mu_{l, 1, i_{\pi(1)}} \cdot ... \cdot \mu_{l, d, i_{\pi(d)}}
    \end{equation*}
    This shows the well-definedness.

    \begin{align*}
        \det(A_I(v_1, ..., v_i + v'_i, ..., v_d)) = \det(A_I(v_1, ..., v_d)) + \det(A_I(v_1, ..., v'_i, ..., v_d))
    \end{align*}
    and
    \begin{align*}
        \det(A_I(v_1, ..., \lambda v_i, ..., v_d)) = \lambda\det(A_I(v_1, ..., v_d))
    \end{align*}
    and
    \begin{equation*}
        \det(A_I(v_1, ..., v_{j + 1}, v_j, ..., v_d)) = -\det(A_I(v_1, ..., v_d))
    \end{equation*}
    However, these properties follow from the well-known properties of the determinant.
    In particular, $\det$ is linear in each column and swapping columns negates the determinant.
    It follows that $f$ is indeed well-defined.
